\documentclass[journal=jacsat,manuscript=article]{achemso}
\setkeys{acs}{articletitle = true, chaptertitle = true}

\usepackage[version=3]{mhchem}
\usepackage{atbegshi}

\author{Wenxi Zhai}
\affiliation{School of Artificial Intelligence and Data Science, University of Science and Technology of China, Hefei 230026, P.R. China}
\affiliation{Suzhou Institute for Advanced Research, University of Science and Technology of China, Suzhou 215123, P.R. China}
\affiliation{Shanghai Innovation Institute, Shanghai 200003, P.R. China}

\author{Jinzhe Zeng}
\email{jinzhe.zeng@ustc.edu.cn}
\affiliation{School of Artificial Intelligence and Data Science, University of Science and Technology of China, Hefei 230026, P.R. China}
\affiliation{Suzhou Institute for Advanced Research, University of Science and Technology of China, Suzhou 215123, P.R. China}
\affiliation{Suzhou Big Data \& AI Research and Engineering Center, Suzhou 215123, P.R. China}

\title{From Tool-Using Language Models to Reliable Scientific Agents for Computational Chemistry}

\begin{document}

\begin{abstract}
Artificial intelligence(AI) is entering a new stage in which models are increasingly embedded in tool-mediated systems that can participate in real-world scientific processes. Yet computational chemistry presents a distinct challenge: successful command execution does not necessarily lead to chemically meaningful, reproducible, or scientifically trustworthy results. In this Perspective, we discuss the transition from tool-using language models to reliable scientific agents for computational chemistry and outline the technical and scientific requirements for verifiable agentic execution. We argue that three requirements are central to this transition. First, the growing complexity of computational chemistry software ecosystems demands skill-based agents that encapsulate reusable scientific procedures rather than merely invoking isolated tools. Second, computational chemistry requires agent frameworks that are more adaptive. Third, because procedural knowledge in computational science is highly task-specific and context-dependent, agentic grounding is essential to connect reasoning with execution traces and error signals. By reframing scientific agents as reliability-oriented execution systems, this Perspective highlights a path toward more rigorous, reproducible, and practically useful agentic computational chemistry.
\end{abstract}

\section{Introduction}
AI systems are increasingly extending beyond standalone models through agentic architectures that connect language-based reasoning with external tools, scientific software, and executable environments. This development broadens the role of AI from producing isolated outputs to coordinating actions within operational frameworks. Recent surveys of LLM-based agents have described this paradigm in terms of tool use, planning, and feedback-driven adaptation. In parallel, agent-skill frameworks treat reusable skills as modular carriers of procedural capability, allowing executable knowledge to be acquired, composed, and updated beyond a single prompt. OpenClaw makes this direction concrete by showing how tasks can be organized through skills, execution contexts, and feedback from real environments. Together, these developments suggest that the next stage of AI agents will depend not only on stronger language understanding, but also on the ability to translate reasoning into structured, tool-mediated, and verifiable action

This transition has quickly entered chemistry and scientific research. Chemistry provides a natural setting for this shift, as chemical discovery requires agents to connect literature knowledge, molecular reasoning, predictive models, synthesis planning, and executable tools. ChemCrow showed that LLMs can be augmented with chemistry-specific tools to support organic synthesis, drug discovery, and materials design, illustrating how general language models can be transformed into chemistry-aware tool-using agents. Recent reviews and perspectives on autonomous agents in chemistry further suggest that the field is moving from isolated tool use toward integrated workflows that combine literature search, synthesis planning, automated experimentation, multimodal reasoning, and closed-loop optimization. Beyond chemistry-specific tasks, Co-Scientist and Robin demonstrate that multi-agent systems can support hypothesis generation, experimental planning, data analysis, candidate ranking, and iterative refinement of scientific hypotheses. In parallel, systems for end-to-end AI research automation indicate that agentic AI is beginning to participate in empirical software development and computational research workflows. In chemistry, the role of LLM agents is being redefined from providers of isolated answers to participants in executable discovery workflows.

However, moving agents into scientific workflows also exposes a deeper challenge. Scientific discovery does not depend only on whether an agent can call a tool, retrieve information, or generate code. It depends on whether the agent can execute a scientifically meaningful process with sufficient reliability, transparency, and reproducibility. This issue is particularly important in computational chemistry, where research workflows often span heterogeneous software ecosystems, complex file formats, and high-performance computing environments. Beyond these engineering challenges, they also depend on domain-specific assumptions, numerical convergence criteria, physical constraints, and extensive post-processing. In this context, successful execution at the command-line level is not equivalent to scientific validity: an apparently completed workflow may still yield chemically unreasonable structures, numerically unreliable results, or outputs that cannot be reproduced because critical provenance information has been lost.

Recent developments around skill-based agents, scientific workflows, and agent safety provide important clues for addressing this gap. OpenClaw-style workflow systems highlight the need to separate general agent control from domain-specific execution skills, so that agents do not merely invoke isolated tools but operate through reusable, inspectable, and task-specific scientific procedures. SkillX further emphasizes the importance of constructing and reusing skill knowledge bases, allowing agents to accumulate procedural knowledge rather than repeatedly solving similar tasks from scratch. Meanwhile, trajectory-level safety and evaluation benchmarks such as ATBench point to the importance of assessing not only final answers, but also the sequence of actions, tool calls, code executions, external effects, and safety-relevant decisions made by agents. These studies collectively suggest that real-world deployment of scientific agents requires more than autonomy: it requires skills, workflows, verification, and safety-aware execution.

In this Perspective, we argue that computational chemistry represents a critical testbed for the next stage of scientific agents. The field already depends on rich computational infrastructures, including quantum chemistry packages, molecular dynamics engines, cheminformatics toolkits, electronic-structure codes, machine-learning potentials, workflow managers, databases, visualization tools, and high-performance computing schedulers. Yet this software complexity also means that tool calling alone is insufficient. Computational chemistry agents must understand how to prepare chemically valid inputs, select appropriate methods, monitor runtime behavior, diagnose failures, validate outputs, estimate uncertainty, preserve provenance, and expose results to human inspection. The next challenge is not whether agents can call computational chemistry tools, but whether they can execute chemically meaningful workflows reliably, transparently, and reproducibly.

We focus on three requirements for this transition from tool calling to reliable execution. First, the increasing complexity of computational chemistry software ecosystems demands skill-based agents that can encapsulate reusable scientific procedures rather than merely invoke isolated tools. Second, computational chemistry requires agent frameworks that are more adaptive than conventional workflows but more reliable than unconstrained LLM agents. Third, because LLMs lack stable, precise, and verifiable procedural knowledge for computational science, agentic grounding is essential to anchor reasoning and decision-making in executable tools, runtime feedback, validation rules, provenance records, and human oversight. By reframing scientific agents as reliability-oriented execution systems, this Perspective outlines what computational chemistry needs from the next generation of scientific agents and how such systems may support trustworthy, reproducible, and practically useful agentic research.

The next challenge is not whether agents can call computational chemistry tools, but whether they can execute chemically meaningful workflows reliably, transparently, and reproducibly.

\section{Software complexity requires skill-based scientific agents}
The increasing complexity of computational chemistry software ecosystems requires scientific agents to move beyond isolated tool access toward skill-based execution. Modern computational chemistry relies on heterogeneous and continuously evolving software infrastructures, including quantum chemistry packages, molecular dynamics engines, density functional theory codes, machine-learning potential frameworks, molecular representation tools, chemical databases, visualization programs, and HPC schedulers. These tools differ in input formats, execution conventions, dependency requirements, runtime environments, and version-specific behaviors. As a result, the practical bottleneck is often not the absence of computational methods, but the difficulty of reliably coordinating software tools, data representations, intermediate files, and execution environments across multistep research tasks.
Simply connecting an LLM agent to more tools does not provide this coordination capability. Tool interfaces expose executable operations, but they rarely encode the expert knowledge needed to use those operations correctly. Computational chemistry workflows involve many implicit and experience-dependent decisions, including molecular preprocessing, unit conversion, charge and spin assignment, multiplicity checking, force-field parameterization, method selection, input validation, convergence diagnosis, failure recovery, and output interpretation. These decisions cannot be fully represented by command-line options or API descriptions alone. The real scientific competence lies in knowing how tools should be combined, under what assumptions they should be used, how intermediate results should be checked, and when execution should be halted or revised.
Skill-based agents provide a more suitable abstraction for this missing layer. A tool exposes a computational capability, whereas a skill encodes task-specific execution competence. In computational chemistry, a skill should not be understood as a simple script. It should package software-facing operations together with usage instructions, workflow templates, parameter constraints, validation checks, expected outputs, and failure-handling logic. In this sense, scientific agents for computational chemistry must move from API-level tool use to workflow-level scientific execution.
The OpenClaw computational chemistry workflow provides a concrete example of this transition. Instead of embedding all domain logic inside a specialized chemistry agent, the framework separates general-purpose agent control from computational chemistry execution. OpenClaw is responsible for coordination, state tracking, skill selection, and supervision. Planning skills externalize natural-language task descriptions into explicit executable task specifications. Domain-specific skills encapsulate computational chemistry procedures and software-facing operations. The DPDispatcher skill further grounds execution in heterogeneous HPC environments by handling job submission, monitoring, and result retrieval. This decoupled design reduces the engineering entanglement among reasoning, workflow specification, domain execution, and infrastructure adaptation.
The associated computational chemistry skill library further illustrates why skillization should be viewed as an ecosystem rather than a static script collection. The current library covers multiple categories, including quantum chemistry, molecular dynamics, machine-learning potentials, molecular representation, atomistic workflows, data processing, analysis, and execution tools. Importantly, new capabilities can be introduced by adding, replacing, or updating domain skills rather than redesigning the core agent. This makes the agent system more maintainable as software packages evolve and new computational methods emerge.
Future scientific agents will likely accumulate capability through the continuous expansion, testing, maintenance, and evolution of such skill ecosystems. A useful computational chemistry skill library should therefore support version control, dependency specification, benchmark tasks, validation tests, provenance tracking, community contribution, and failure-case accumulation. Under this view, the long-term progress of scientific agents will not depend only on larger language models or longer tool lists, but on whether expert knowledge can be encoded into reusable, testable, auditable, and evolvable skills.

\section{Reliable agents should be more adaptive than workflows and more constrained than general LLM agents}
Computational chemistry requires an automation framework that is more flexible than conventional workflows but more reliable than unconstrained LLM agents. Traditional workflow systems have been crucial for reproducible and traceable computation. By defining task dependencies, organizing data flow, recording provenance, and connecting calculations to computing resources, they provide a robust foundation for repeated and well-specified procedures. However, many computational chemistry tasks cannot be fully reduced to a fixed task graph in advance. Cross-software adaptation, runtime errors, context-dependent decisions, and unexpected intermediate results often require the workflow to be inspected, revised, or partially recovered during execution.

General LLM agents provide complementary flexibility. They can interpret natural-language goals, decompose tasks, select tools, generate scripts, and adapt plans based on intermediate outputs. Yet this flexibility can become a liability when planning logic, tool routing, recovery strategy, and execution assumptions are hidden inside prompts or specialized agent stacks. In such systems, it is often difficult to determine which part of the agent controls the workflow, which part enforces domain constraints, and which part is responsible for recovering from failure. This makes the system harder to inspect, maintain, and validate.

A useful computational chemistry agent should therefore occupy a middle ground. It should preserve the agent’s capacity for dynamic planning while making workflows, states, intermediate results, and recovery actions explicit. Tools execute commands; skills execute scientific intentions. The key is not to replace workflows with open-ended autonomy, but to make workflows more adaptive, state-aware, and recoverable under scientific constraints.

The OpenClaw computational chemistry workflow provides a concrete example of this intermediate paradigm. Schema-defined planning skills convert natural-language scientific goals into executable task specifications, rather than leaving the entire plan implicit in the model prompt. The task manifest then specifies which skills should be combined, in what order, under what dependencies, and with what completion or validation conditions. This makes the workflow more inspectable than a purely prompt-driven agent while remaining more adaptable than a fully pre-encoded workflow.

The execution layer is equally important. Through the DPDispatcher skill, job submission, status monitoring, result retrieval, and heterogeneous HPC scheduling are treated as manageable workflow states rather than ad hoc shell operations. In the methane oxidation reactive MD case study, this design enabled the agent to coordinate molecular preparation, geometry optimization, file conversion, reactive system construction, molecular dynamics simulation, HPC dispatch, runtime monitoring, failure recovery, and trajectory analysis. The case also showed bounded recovery from execution failures and downstream reaction-network extraction. These results illustrate why computational chemistry needs a framework that combines the flexibility of agents with the explicitness, traceability, and recovery semantics of workflows.


\section{Agentic grounding for procedural scientific knowledge}

LLMs can integrate textual knowledge, generate plans, write scripts, and explain results, but they do not inherently possess stable, precise, and verifiable procedural knowledge for computational science. Computational chemistry requires more than knowing the names of methods or software packages. It requires correct file formats, parameter conventions, software-specific execution rules, HPC environments, numerical checks, physical constraints, and expert judgment about whether a result is meaningful. Without grounding in these executable and verifiable elements, an LLM may produce a plausible workflow that fails in practice, or worse, a workflow that runs successfully but leads to unreliable scientific conclusions.

This limitation motivates the need for agentic grounding. In computational chemistry, agentic grounding means anchoring the reasoning and decisions of an LLM-based agent in executable tools, domain-specific skills, runtime feedback, validation rules, provenance records, and human oversight. The goal is not to make the LLM an independent scientist that internally stores all computational chemistry expertise. Rather, the goal is to place the LLM within a constrained scientific execution system, where its planning ability is continuously checked by real software behavior, explicit task specifications, intermediate outputs, error messages, and domain validation criteria.

The OpenClaw computational chemistry workflow illustrates this principle. OpenClaw acts as a general control layer for coordination and supervision, rather than encoding all computational chemistry logic directly in the model or the core agent. Planning skills and domain skills provide explicit procedural constraints: they translate broad scientific goals into executable task specifications, define how software operations should be combined, and expose the conditions under which execution can proceed. In this architecture, the LLM reasons over the user goal, loaded skills, current state, and execution feedback, while OpenClaw performs actions, captures tool outputs, and feeds them back into the control loop.

Runtime feedback is central to this grounding process. Logs, return codes, generated files, schema checks, intermediate artifacts, and scheduler states provide evidence that constrains subsequent reasoning. When execution fails, the agent does not simply continue generating text; it must inspect the failure signal, select a bounded recovery action, request human guidance, or halt the workflow. This converts software errors and environmental constraints into part of the agent’s decision-making context. As a result, model reasoning becomes coupled to the actual behavior of computational tools rather than remaining at the level of abstract planning.

The methane oxidation reactive MD case further shows that the value of such an agent is not to replace chemical knowledge, but to coordinate a complex computational process. The system organized molecular preparation, geometry optimization, file conversion, reactive system construction, molecular dynamics simulation, HPC dispatch, failure recovery, and trajectory analysis. Its contribution was procedural: maintaining workflow state, coordinating cross-tool execution, recovering from runtime failures within defined limits, and producing traceable outputs for downstream expert assessment. Scientific validity still depends on human domain experts who choose physical models, inspect inputs, evaluate numerical protocols, and interpret final observables.

Therefore, reliable scientific agents for computational chemistry should be understood as grounded execution systems rather than autonomous scientific authorities. Their strength lies in connecting LLM reasoning with executable procedures, software feedback, validation mechanisms, and expert supervision. This perspective reframes agentic computational chemistry: the central challenge is not to make LLMs memorize more procedural knowledge, but to design systems in which procedural knowledge is externalized, tested, updated, and enforced through the interaction between agents, skills, tools, environments, and scientists.


\section{OpenClaw-grounded computational chemistry automation as an integrated case}

The OpenClaw-based computational chemistry workflow provides an integrated example of how reliable execution-oriented agents may be built for complex scientific tasks. Its central contribution is not simply to connect an LLM with computational chemistry tools, but to propose a decoupled agent–skill design in which general-purpose agent control is separated from domain-specific scientific execution. Under this design, OpenClaw serves as the centralized control and supervision layer, responsible for context management, skill selection, task coordination, state tracking, and execution supervision. Computational chemistry knowledge and software-facing procedures are externalized into reusable skills rather than being embedded entirely in the model prompt or hard-coded into a specialized agent.

This architecture contains three key components. First, schema-defined planning skills translate natural-language scientific goals into executable task specifications. Instead of leaving the workflow implicit in LLM reasoning, the system constructs a task manifest that defines stages, dependencies, required inputs, expected outputs, validation conditions, and execution constraints. Second, domain skills encapsulate computational chemistry procedures, including software-specific operations, templates, checks, and result-handling logic. These skills provide the procedural layer needed to turn scientific intentions into executable operations. Third, the DPDispatcher skill grounds the workflow in heterogeneous HPC environments by handling job script generation, job submission, status monitoring, and result retrieval across local and remote computing systems.

The methane oxidation reactive molecular dynamics case demonstrates how this design works in practice. Starting from a natural-language instruction, the system organized a multistage workflow involving molecular preparation, quantum-chemical geometry optimization, file format conversion, reactive system construction, molecular dynamics simulation, HPC dispatch and monitoring, runtime failure recovery, and trajectory analysis. The workflow coordinated multiple tools, including Open Babel, Gaussian, dpdata, Packmol, LAMMPS, DeePMD-kit, DPDispatcher, and ReacNetGenerator. Importantly, the task was not executed as a fixed linear script. Progression between stages depended on intermediate outputs, validation checks, scheduler feedback, and bounded recovery actions when failures occurred.

This case is significant because it shows a maintainable and extensible route toward multistep computational chemistry automation. New capabilities can be added by updating planning or domain skills rather than redesigning the core agent. Execution can be made more traceable through task manifests, generated files, logs, scheduler states, and intermediate artifacts. Failures can be handled within defined recovery boundaries instead of being hidden inside uncontrolled prompt-based behavior. At the same time, the case should not be interpreted as a demonstration of fully autonomous scientific discovery. The agent coordinates and executes a complex computational process, but the assessment of physical models, numerical protocols, input quality, sampling adequacy, and mechanistic interpretation still requires human domain expertise.

Thus, the OpenClaw case provides a concrete bridge between tool-using LLM agents and reliable scientific execution systems. It illustrates that practical computational chemistry agents require more than tool access: they require explicit task specifications, reusable domain skills, grounded execution environments, runtime feedback, bounded recovery, provenance, and human validation. In this sense, OpenClaw is best understood as an early example of how agentic computational chemistry can become maintainable, extensible, and scientifically accountable without claiming to replace expert scientific judgment.


\section{Outlook}
Looking forward, the development of reliable scientific agents for computational chemistry should move beyond simply expanding tool access and instead focus on building trustworthy execution infrastructure. Key challenges include quality control of skills, clear documentation of their sources, versions, dependencies, assumptions, applicability boundaries, tests, and validation criteria; the construction of self-evolving skill ecosystems that can extract, correct, and update procedural knowledge from historical tasks and failure traces; and rigorous provenance tracking, audit trails, environment records, and reproducibility mechanisms. At the same time, successful execution must not be equated with scientific correctness: physical models, numerical protocols, sampling adequacy, uncertainty, and downstream interpretation still require domain expertise. As agents gain access to files, software, databases, HPC resources, and potentially experimental platforms, safety and governance mechanisms such as permission control, sandboxed execution, trajectory-level evaluation, and human oversight will become increasingly important. The OpenClaw computational chemistry workflow provides an early example of how a general-purpose agent, domain-specific skills, and an HPC execution layer can support maintainable and extensible multistep automation, but its significance lies in reliable coordination rather than fully autonomous discovery. The next stage of agentic computational chemistry should therefore aim to connect simulation, experiment, data analysis, and knowledge updating into verifiable closed loops, while ensuring that scientific agents remain reproducible, auditable, maintainable, and aligned with expert human judgment.

\begin{acknowledgement}

acknowledgement

\end{acknowledgement}
% Uncomment this after adding at least one \cite{...} in the text.
% \bibliography{combined}

\end{document}
